Towards Compiler-Agnostic Performance in Finite-Difference Codes. (TBC)

### 2014-11-21 shallow_perf.tex
Add text to Intro and Conclusions

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HE" in caps to complete the first word.
\IEEEPARstart{T}{he} challenge presented to the developers of
scientific software by the drive towards Exascale computing is
considerable. With power consumption becoming the overriding design
constraint, CPU clock speeds are falling and the complex,
multi-purpose compute core is being replaced multiple, simpler
cores. This philosophy can be seen at work in the rise of so-called
accelerator based machines in the Top 500 List~\cite{top500} of
supercomputers: five of the top-ten machines in the November 2014 list
make use of Intel Xeon Phi's or NVIDIA GPUs. Four of the remaining
five machines in the top ten are IBM BlueGene/Qs, the CPU of which has
hardware support for running 64 threads.

@@

In fact, for the Gnu and Cray compilers on the Ivy Bridge system, the
\psykal version of shallow is never more than 2\% slower than the
original and in some cases is faster.  This demonstrates that we can
reliably recover the performance of the original version of the code,
despite the significant restructuring required by the \psykal
approach. {\bf We need to re-run/check the Intel results on Ivy Bridge
  once Archer is back from its upgrade.}

@@

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

@@

\section{Conclusions}
The conclusion goes here.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HE" in caps to complete the first word.
\IEEEPARstart{T}{he} challenge presented to the developers of
scientific software by the drive towards Exascale computing is
considerable. With power consumption becoming the overriding design
constraint, CPU clock speeds are falling and the complex,
multi-purpose compute core is being replaced by multiple, simpler
cores. This philosophy can be seen at work in the rise of so-called
accelerator based machines in the Top 500 List~\cite{top500} of
supercomputers: five of the top-ten machines in the November 2014 list
make use of Intel Xeon Phi's or NVIDIA GPUs. Four of the remaining
five machines in the top ten are IBM BlueGene/Qs, the CPU of which has
hardware support for running 64 threads.

Achieving good performance on large numbers of light-weight cores
requires exploiting as much parallelism in an application as possible
and this results in increased complexity in the programming models
that must be used. This in turn increases the burden of code
maintenance and code development, in part because two specialisms are
required: that of the scientific domain which a code is modelling
({\it e.g.} oceanography) and that of computational science. The
situation is currently complicated still further by the existence of
competing hardware technology; if one was to begin writing a major
scientific application today it is unclear whether one would target
GPU, Xeon Phi, traditional CPU, FPGA or something else entirely. This
is a problem because, generally speaking, these different technologies
require different programming approaches.

@@

In fact, for the Gnu and Cray compilers on the Ivy Bridge system, the
\psykal version of shallow is never more than 2\% slower than the
original and in some cases is faster.  This demonstrates that we can
reliably recover the performance of the original version of the code,
despite the significant restructuring required by the \psykal
approach.

@@

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.

%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

@@

\section{Conclusions}

We have investigated the application of the \psykal separation of
concerns approach to the domain of finite-difference ocean
models. This approach enables the computational science (performance)
related aspects of a computer model to be kept separate from the
natural (oceanographic) science aspects. In this work we have shown
that the \psykal approach has the potential to improve performance and
to improve code structure/maintainability.

The application of code transformations to the middle/PSy layer is key
to the performance of the \psykal version of a code.
